---
title: "ARE 212 Problem Set 3"
author: "Eleanor Adachi, Aline Adayo, Anna Cheyette, Karla Neri, and Stephen Stack"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  - \usepackage[T1]{fontenc}
  - \usepackage{textcomp}
  - \usepackage{lmodern}
  - \usepackage{underscore}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set up}
# Comment out after installing
# install.packages("pacman")

options(scipen = 999)

# Load packages
library(pacman)
# p_load(dplyr, haven, readr, knitr, psych, ggplot2,stats4, stargazer, lmSupport, 
#        magrittr, qwraps2, Jmisc , fastDummies)

# Remove lmSupport, add tidyverse
p_load(dplyr, haven, readr, knitr, psych, ggplot2,stats4, stargazer, magrittr, 
       qwraps2, Jmisc , fastDummies, tidyverse)

# get directory of current file
current_directory <-
  dirname(dirname(rstudioapi::getSourceEditorContext()$path))
```

# Question 1


*Read* `pset3_2024.dta` *into R Please check for missing values (as in section 3).*

```{r load data}
# Load data
my_data_raw <- read_dta(file.path(current_directory, "data", "pset3_2024.dta"))
head(my_data_raw)

# Check for missing values
my_data <- my_data_raw %>% drop_na()

# Check if any missing values removed
all.equal(my_data, my_data_raw)
```

# Question 2

*Get the summary statistics for price: sample mean, standard deviation, minimum and maximum. Construct a 99% confidence interval for the sample average of price (car price in thousands of Euros).*

```{r summary statistics for price}
# Get summary statistics for price
describe(my_data$price)
```

```{r summarize data, results='asis'}
# Create summary table
summary_maker <-
  list("Price" =
         list("min" = ~ min(my_data$price),
              "max" = ~ max(my_data$price),
              "mean (sd)" = ~ qwraps2::mean_sd(my_data$price)))

whole <- summary_table(my_data, summary_maker)
whole
```

```{r confidence interval, results='asis'}
# Construct a 99% confidence interval for the sample average of price
xbar <- mean(my_data$price)
n <- nrow(my_data)
se_xbar <- sd(my_data$price)/sqrt(n)
cnn <- qt(0.005,n-1,lower.tail=FALSE)
bottom99ci <- xbar-cnn*se_xbar
top99ci <- xbar+cnn*se_xbar
bottom99ci
top99ci
```

The 99% confidence interval for the sample mean of price is **`r bottom99ci`** to **`r top99ci`**.

# Question 3

*Create two new variables log of price and log of quantity,* `lprice` *and* `lqu` *Create the scatter plot of the two variables* `lqu` *and* `lprice`. *What is the estimated OLS linear model slope associated with this scatter plot? Estimate a regression to answer this.*

```{r create new variables}
# Create lprice and lqu
my_data$lprice <- log(my_data$price)
my_data$lqu <- log(my_data$qu)
```

```{r create lprice-lqu scatterplot}
# create lprice-lqu scatterplot
X3 <- my_data$lprice
y3 <- my_data$lqu
lscatter <- ggplot() +
  geom_point(aes(x = X3, y = y3)) +
  labs(x = "Log of Car Price (in 1000 Euros)",
       y = "Log of Car Sales Quantity",
       title = str_wrap( 
         "Log of Car Sales Quantity vs. Log of Car Price", 
         40)) +
  ylim(0,max(y3)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
lscatter
```

```{r estimate slope using OLS}
# TODO: Should we be using a constant in OLS regression??? ####

# estimate slope using OLS
ols_estimator <-function(y,X){
  # Sample size
  n <- length(y)
  # degrees of freedom
  if (is.null(dim(X))) {
    df <- n - 1
    } else {
      df <- nrow(X) - ncol(X)
      }
  # Find coefficient vector
  b <- solve(t(X) %*% X) %*% t(X) %*% y
  # projection matrix of reg y on X
  P <- X%*%solve(t(X)%*%X)%*%t(X)
  # residual maker of reg y1 on x: M= I - P
  M <- diag(n)-P
  # sum of squared residuals, SSR=e'e
  e <- M%*%y
  SSR <- t(e)%*%e
  # construct demeaner
  i <- c(rep(1,n))
  M0 <- diag(n)-i%*%t(i)*(1/n)
  # demeaned y
  M0y <- M0%*%y
  # total sum of squares
  SST <- t(M0y)%*%M0y
  # calculate R squared
  Rsquared <- 1-(SSR/SST)
  # create predictions
  y_hat <- P%*%y
  #varcov matrix b
  s2 <- as.numeric(t(e)%*%e)/df
  vb <- s2*solve(t(X)%*%X)
  #std error b
  seb <- sqrt(diag(vb))
  return(list(b, Rsquared, y_hat, e, seb))
}

results3 <- ols_estimator(y3, X3)
b3 <- results3[[1]]
b3
```

The estimated slope of this regression using OLS (without a constant) is **`r b3`**.

# Question 4

*Regress lqu on fuel, luxury, domestic, and a constant, create the residuals elqu.*

```{r create elqu}
# Regress lqu on fuel, luxury, domestic, and a constant
X4 <- cbind(1, my_data$fuel, my_data$luxury, my_data$domestic)
y4_lqu <- my_data$lqu
results4_lqu <- ols_estimator(y4_lqu, X4)

# create the residuals elqu
elqu <- results4_lqu[[4]]
```

*Regress lprice on fuel, luxury, domestic, and a constant, create the residuals elprice*

```{r create elprice}
# Regress lprice on fuel, luxury, domestic, and a constant
y4_lprice <- my_data$lprice
results4_lprice <- ols_estimator(y4_lprice, X4)

# create the residuals elprice
elprice <- results4_lprice[[4]]
```

*Scatter plot the residuals elqu on vertical axis and elprice on horizontal axis*

```{r scatter plot elqu vs. elprice}
# Scatter plot the residuals elqu on vertical axis and elprice on horizontal axis
ggplot() +
  geom_point(aes(x = elprice, y = elqu)) +
  labs(x = "Residuals of Log of Car Price",
       y = "Residuals of Log of Car Sales Quantity",
       title = "Log Quantity Residuals vs. Log Price Residuals"
       ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```
*What is the estimated OLS slope associated with this scatter plot? Estimate a regression (no constant) to answer this and explain what theorem underlies the fact that this slope is the marginal effect of lprice on lqu in a regression that also features fuel, luxury, domestic, and a constant.*

```{r OLS of scatter plot elqu vs. elprice}
results4_elqu <- ols_estimator(elqu, elprice)
b4_elqu <- results4_elqu[[1]]
b4_elqu
```

The OLS slope for the residuals `elqu` on the residuals `elprice`, **`r b4_elqu`**, is equal to the coefficient of `lprice` in the regression of `lqu` on `lprice`, fuel, luxury, domestic, and a constant. This is a demonstration of the Frish-Waugh-Lovell Theorem.

# Question 5

*Why is the slope estimate in 3 not equal to the one in 4? Theoretically speaking, when would they be equal?*

The slope estimates in 3 and 4 are not equal because there is some relationship between `lprice` and the other variables in 4. The estimates would be equal if the vectors were orthogonal, i.e., there was no relationship between `lprice` and the other variables.

# Question 6

*Please interpret the OLS slope point estimate size, sign of the slope lprice estimate in 4.*

The OLS slope for the residuals `elqu` on the residuals `elprice`, **`r b4_elqu`**, means that for each 100% increase in `price` there is a ~333% decrease in `qu`. We know that it is a decrease because the slope is negative.

*What is the pvalue for the estimated lprice coefficient? Use the stat tables for this. And then check with* $pvalue_6 = 2 \cdot pt(abs(t_6),df)$ *, where t6 is the t stat value t6=-, and df are degrees of freedom*

```{r pvalue for lprice coefficient}
# pvalue for lprice coefficient
seb6_elqu <- results4_elqu[[5]]
t6 <- b4_elqu/seb6_elqu
df6 <- length(elqu) - 1 # only one parameter
t6
```

The $t$ statistic is **`r t6`** and there are **`r df6`** degrees of freedom. Looking at the t-tables, this corresponds to a p-value of less than 0.000.

```{r check pvalue}
# pvalue for lprice coefficient
pvalue6 <- 2*(1-pt(abs(t6),df6))
pvalue6
```

The p-value is **`r pvalue6`**.


# Question 7

*Can you reject that the marginal effect of lprice on lqu is -4 conditional on all else equal (fuel, luxury, domestic, and a constant)? Do five steps in Hypothesis Testing at the 5% significance level against a two-sided alternative. Get critical values from the relevant stats table.*

$H_0: \beta = -4$
$H_a: \beta \neq -4$

Remember, estimate of $\beta$ is *b*, **`r b4_elqu`** and standard error is **`r seb6_elqu`**.

```{r find t}
# get t statistic
t7 <- (b4_elqu - (-4))/seb6_elqu
```

Critical value of *t* for 5% significance and **`r df6`** degrees of freedom from table is between 2.000 and 2.021.

The *t* statistic is **`r t7`**, which is less than the critical value. Therefore, we cannot reject the null hypothesis.

# Question 8

*Estimate the sample data correlation of all these variables with each other: lqu, lprice, fuel, weight, luxury, domestic. Suppose the population model is given by*
$lqu=beta0+ lprice beta1 + domestic beta2 +fuel beta3 + luxury beta4 +epsilon (8.a)$
*and you estimate the model*
$lqu=alpha0+ lprice alpha1 + fuel alpha3 +luxury alpha4 +epsilon (8.b)$

*Based on the variables’ correlation and without estimating any regression models, would the estimated coefficient for fuel in (8.b) have a negative or a positive bias? Explain briefly.*

```{r find correlations}
# correlation between fuel and domestic
cor(my_data$fuel,my_data$domestic)

# correlation between luxury and domestic
cor(my_data$luxury,my_data$domestic)

# correlation between lprice and domestic
cor(my_data$lprice,my_data$domestic)
```
Domestic cars tend to use more fuel and more fuel-intensive cars tend to be domestic.

The estimated coefficient for fuel in (8.b) would have a negative bias.

Since fuel and domestic are closely related, the coefficients on a regression that includes domestic would tend to be more inflated (positive bias). Therefore the regression that doesn't include domestic would tend to have "deflated" (negative bias) effect on lqu.

```{r run regression for 8.b}
# Regress lqu on lprice, fuel, luxury, and a constant
X8b <- cbind(1, my_data$lprice, my_data$fuel, my_data$luxury)
y8b <- my_data$lqu
results8b <- ols_estimator(y8b, X8b)

# Return coefficients
b8b <- results8b[[1]]
b8b
```

# Question 9

*If I told you that research shows that advertising expenditures by car model are positively correlated with lprice and that when including advertising in addition to all factors in (8.b), the estimated weight coefficient does not change at all. What does this imply about the sample correlation between advertising and weight of cars in the sample?*

The sample correlation between advertising expenditures and car weight is very weak.

# Question 10

*Suppose that research showed that the log of advertising is, on average, 5 times the log of price. Construct that advertising variable based on this fact and include it in a regression in addition to lprice and the other covariates in 8.b.*

```{r run regression with advertising}
# Regress lqu on lprice, fuel, luxury, ladvertising and a constant
my_data$ladvertising <- 5*my_data$lprice
X10 <- cbind(1, my_data$lprice, my_data$fuel, my_data$luxury, my_data$ladvertising)
y10 <- my_data$lqu

# Commented out to knit document because returns error
#results10 <- ols_estimator(y10, X10)
```

*Explain what happened.*

This returns an error:

Error in solve.default(t(X) %*% X) : system is computationally singular: reciprocal condition number = 5.09482e-34

Since one of the explanatory variables is a linear combination of another explanatory variable, this violates one of the conditions of OLS.

# Question 11

*Please estimate a specification that allows you to test the following. Research shows that luxury goods have a different price elasticity than non-luxury goods. The null hypothesis is that the marginal effect in lprice on lqu does not differ by luxury classification of the car. Write out the regression model that allows you to estimate and perform a hypothesis test for this null. Do the five steps in hypothesis testing at the 5% significance level. What do you conclude?*

```{r test null hypotehsis that luxury does not affect coefficient of lprice}
# Regress lqu on lprice, luxury, interaction term, and a constant
reg11 <- lm(lqu~lprice + luxury + lprice*luxury, my_data)
summary(reg11)

# Find p-value
pval11 <- summary(reg11)$coefficients[4, 4]
```

Null hypothesis: $H_0$: $\beta_3 = 0$

Alternative hypothesis: $H_a$: $\beta_3 \neq 0$

Since the p-value of the null hypothesis is **`r pval11`** which is greater than 0.05, we cannot reject the null hypothesis.

We cannot conclude that the marginal effect of `lprice` on `lqu` differs between luxury and non-luxury cars.


# Question 12

*Regress lqu on a constant, fuel, lprice, luxury, domestic, weight. (eq 12) Test the joint hypothesis that* $\beta_{domestic}=1.5$ *and* $\beta_{fuel}=60 \beta_{weight}$ *at the 1 percent significance level.*

```{r regress lqu on a constant, fuel, lprice, luxury, domestic, weight}
# Regress lqu on a constant, fuel, lprice, luxury, domestic, weight
X12 <- cbind(1, my_data$fuel, my_data$lprice, my_data$luxury, my_data$domestic, 
             my_data$weight)
y12 <- my_data$lqu
results12 <- ols_estimator(y12, X12)
b12 <- results12[[1]]
e12 <- results12[[4]]
df12 <- nrow(X12) - ncol(X12)
s2_12 <- as.numeric(t(e12)%*%e12)/df12
```

*a) Perform a Fit based test and then also a Wald test.*

```{r fit based test}
# Test joint hypothesis that beta_domestic=1.5 and beta_fuel=60*beta_weight
# F = (ssrr-ssr)/(2*(s2))
# ssr restricted
X12r <- cbind(1, (60*my_data$fuel + my_data$weight), my_data$lprice, 
              my_data$luxury)
y12r <- my_data$lqu - 1.5*my_data$domestic
results12r <- ols_estimator(y12r, X12r)
e12r <- results12r[[4]]
ssrr <- as.numeric(t(e12r)%*%e12r)

# ssr unrestricted
ssru <- as.numeric(t(e12)%*%e12)
# F test
F_fitBased<-(ssrr-ssru)/(2*(s2_12))
F_fitBased
```

```{r Wald test}
# Test joint hypothesis that beta_domestic=1.5 and beta_fuel=60*beta_weight
Rr1 <- c(0, 0, 0, 0, 1, 0)
Rr2 <- c(0, 1, 0, 0, 0, -60)
R <- t(cbind(Rr1,Rr2))
q<-c(1.5, 0)
VRbq <- s2_12* R %*% solve(t(X12) %*% X12) %*% t(R)
Fw <- (t(R %*% b12-q) %*% solve(VRbq) %*% (R %*% b12-q))/2 # J = 2
Fw
```

*b) Are the values of the fit and Wald test statistics equal?*

Yes, the F-statistic values of the fit-based F test and the Wald test are equal.


# Question 13

*Without running any additional regressions and starting from the baseline regression in reg question 8.a,*

*a) Will omitting fuel create an OVB problem for the OLS estimator of lprice?*

Population model given by (8.a):

$lqu = \beta_0 + lprice \beta_1 + domestic \beta_2 + fuel \beta_3 + luxury \beta_4 + \epsilon$

Sample model if fuel omitted:

$\hat{lqu} = \tilde{\beta_0} + lprice \tilde{\beta_1} + domestic \tilde{\beta_2} + luxury \tilde{\beta_4} + \tilde{e}$

```{r correlation between fuel and lqu}
# correlation between fuel and lqu
cor(my_data$fuel,my_data$lqu)
```
Increasing fuel efficiency tends to decrease quantity sold.

```{r correlations between other explanatory variables and fuel}
# correlation between lprice and fuel
cor(my_data$lprice,my_data$fuel)

# correlation between domestic and fuel
cor(my_data$domestic,my_data$fuel)

# correlation between luxury and fuel
cor(my_data$luxury,my_data$fuel)
```
Increasing price, increasing domestic (true = 1), and increasing luxury (true = 1) tends to increase fuel efficiency.

Since the coefficients have opposite signs, omitting fuel will lead to omitted variable bias and the bias will be negative. The coefficients for the other explanatory variables will tend to be underestimated.

*b) Compute the variance inflated factor (VIF) for the variable height to be potentially also included into the (reg of question 8.a) model to explain the variation in lqu. Feel free to use the lm canned function to get what you need for the VIFj, for all j.*

```{r compute VIF for height}
# regress height on other explanatory variables
reg13 <- lm(height~lprice+domestic+fuel+luxury, my_data)
summary(reg13)
rsq13 <- summary(reg13)$r.squared
vif13 <- 1/(1-rsq13)
vif13
```
The variance inflated factor (VIF) for height is **`r vif13`**.

*c) Will including this variable height with the others in (model in question 8.a) result in multicollinearity problems?*

There is not strong evidence that including height in the model from 8.a will results in multicollinearity problems. The R-squared value is only 0.07552 and the VIF is **`r vif13`**, which is only slightly more than 1.

# Question 14

*Suppose a car salesman told you that the conditional variance in the unobserved determinants of the log quantity (lqu) for luxury cars is three times the variance for nonluxury cars.*

*a. Which assumption no longer holds when we derive the statistical properties of the OLS estimators for the linear model in (reg 8.a) ?*

Assumption 3--homoskedasticity and spherical disturbances--no longer holds.

*b. Let the variance of the disturbance of log quantity for luxury=1 be 3 times the variance for luxury=0. In R create a matrix Omega, its inverse, and the positive definite matrix C such that the inverse of Omega = C C’ as derived in lecture.*


```{r create Omega, inverse of Omega, and positive definite matrix C}
# sort data by luxury
my_dataGLS <- my_data[order(my_data$luxury),]
head(my_dataGLS)

# Count number of non-luxury and luxury
countlux <- sum(my_dataGLS$luxury == 1)
countnonlux <- nrow(my_dataGLS) - countlux

# Construct Omega: 54 non-luxury (luxury = 0), 3 luxury (luxury = 1)
topnonlux <- diag(countnonlux)
toplux <- matrix(data=0, nrow=countnonlux, ncol=countlux)
topOmega<-cbind(topnonlux, toplux)
bottomnonlux <- matrix(data=0, nrow=countlux, ncol=countnonlux)
bottomlux <- diag(countlux)*3 #identity times 3
bottomOmega <- cbind(bottomnonlux, bottomlux)
Omega <- rbind(topOmega,bottomOmega)

# Calculate inverse of Omega
invOmega <- solve(Omega)

# Calculate positive definite matrix C, sqrt of invOmega
invOm.eig <- eigen(invOmega)

D <- diag((invOm.eig$values))
V <- invOm.eig$vectors
C <- invOm.eig$vectors %*% diag(sqrt(invOm.eig$values)) %*% solve(invOm.eig$vectors)
```

*c. Estimate the BLUE estimator in this setting of model 8.4 and test whether the marginal effect of lprice on lqu is equal to -1, at the 10 percent significance level, ceteris paribus.*

Population model given by (8.a):

$lqu = \beta_0 + lprice \beta_1 + domestic \beta_2 + fuel \beta_3 + luxury \beta_4 + \epsilon$

```{r estimate BLUE}
# pre multiply all the variables by C, get ytilde and Xtilde
ytilde <- C %*% my_dataGLS$lqu
Xtilde <- C %*% cbind(1, my_dataGLS$lprice, my_dataGLS$domestic, my_dataGLS$fuel, 
                      my_dataGLS$luxury)

bGLS <- solve(t(Xtilde)%*%Xtilde) %*% t(Xtilde)%*%ytilde
eGLS <- ytilde - Xtilde%*%bGLS

S2_GLS <- (t(eGLS) %*% eGLS)/(length(eGLS)-ncol(Xtilde))
S2_GLS <- as.numeric(S2_GLS)

# variance of bGLS
VbGLS <- solve(t(Xtilde)%*%Xtilde) *  S2_GLS

# standard error of bGLS
seGLS <- sqrt(diag(VbGLS))
```

```{r use BLUE to test whether beta_lprice is -1 at 10% significance level}
t14 <- (bGLS[[2]] - (-1))/seGLS[[2]]
t14

df14 <- nrow(Xtilde) - ncol(Xtilde)
tc14 <- qt(.05, df14, lower.tail=TRUE)
tc14
```
Null hypothesis: $H_0 : \beta_1 = -1$

Alternative hypothesis: $H_a : \beta_1 \neq -1$

The t score for the `lprice` coefficient, $\beta_1$, being equal to -1 is **`r t14`**. The critical value for the 10% significance level is **`r tc14`**. Since $|t| > |t_{critical}|$, we reject the null hypothesis and conclude that $\beta_1$ is not equal to -1.

